\Rpackage{MSnbase}: labelled MS2 data pre-processing, visualisation and quantification.

### 2012-12-14 inst/doc/MSnbase-demo/MSnbase-demo.Rnw

\newcommand{\R}{\texttt{R} }
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
%% colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}

@@

<<knitr, include=FALSE, cache=FALSE>>=
library(knitr)
opts_chunk$set(fig.align = 'center', 
               fig.show = 'hold', 
               par = TRUE,
               prompt = TRUE,
               comment = NA)
options(replace.assign = TRUE, 
        width = 68)
knit_hooks$set(par = function(before, options, envir) {
  if (before && options$fig.show != 'none') 
    par(mar = c(4,4,.1,.1),
        cex.lab = .95,
        cex.axis = .9,
        mgp = c(2,.7,0),
        tcl = -.3)
})
@

@@

If no peak is detected for a reporter ion peak, the respective 
quantitation value is set to \texttt{NA}. In our case, there 
is \Sexpr{sum(is.na(exprs(qnt)))} such case in row 
\Sexpr{which(is.na(exprs(qnt))) %% nrow(qnt)}.
We will remove the offending line, as we do not expect any missing peaks.

<<removeNa, echo=TRUE>>=
table(is.na(qnt))
whichRow <- which(is.na((qnt))) %% nrow(qnt)
qnt <- qnt[-whichRow, ]
sum(is.na(exprs(qnt)))
@

The \Rfunction{filterNA} method allows to filter features based on the proportion 
of missing data their carry, specified using the \Robject{pNA} argument. 
The same result as above could have been achieved with 
\Rfunction{qnt <- filterNA(qnt, pNA = 0)}, 
specifying that we allow 0 percent of missing data. 
See also the \Rfunction{plotNA} method to get an overview of the completeness 
of a data set.

@@

<<make2exps2, echo=TRUE, cache=FALSE>>=
set.seed(1)
i <- sample(length(itraqdata), 35)
j <- sample(length(itraqdata), 35)
exp1 <- quantify(itraqdata[i], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
exp2 <- quantify(itraqdata[j], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
table(featureNames(exp1) %in% featureNames(exp2))
exp1 <- combineFeatures(exp1,
                        groupBy = fData(exp1)$ProteinAccession)
exp2 <- combineFeatures(exp2,
                        groupBy = fData(exp2)$ProteinAccession)
head(featureNames(exp1))
head(featureNames(exp2))
@

@@

>>>>

\newcommand{\R}{\texttt{R} }
\newcommand{\code}[1]{{\texttt{#1}}}
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\mbox{\normalfont\textsf{#1}}}}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}
%% colors
\definecolor{Red}{rgb}{0.7,0,0}
\definecolor{Blue}{rgb}{0,0,0.8}

@@

<<knitr, include=FALSE, cache=FALSE>>=
library(knitr)
opts_chunk$set(fig.align = 'center', 
               fig.show = 'hold', 
               par = TRUE,
               prompt = TRUE,
               comment = NA)
options(replace.assign = TRUE, 
        width = 68)
## see ?evaluate::evaluate for details
opts_knit$set(stop_on_error = 1L)

## knit_hooks$set(par = function(before, options, envir) {
##   if (before && options$fig.show != 'none') 
##     par(mar = c(4,4,.1,.1),
##         cex.lab = .95,
##         cex.axis = .9,
##         mgp = c(2,.7,0),
##         tcl = -.3)
## })
@

@@

If no peak is detected for a reporter ion peak, the respective 
quantitation value is set to \texttt{NA}. In our case, there 
is \Sexpr{sum(is.na(exprs(qnt)))} such case in row 
\Sexpr{which(is.na(exprs(qnt))) %% nrow(qnt)}.
We will remove the offending line using the \Rfunction{filterNA} method.
The \Robject{pNA} argument defines the percentage of accepted missing 
values per feature. As we do not expect any missing peaks, we set it 
to be 0 (which is also the detault value).

<<filterNA, echo=TRUE>>=
table(is.na(qnt))
qnt <- filterNA(qnt, pNA = 0)
sum(is.na(qnt))
@

The infrastructure around the \Robject{MSnSet} class allows flexible filtering 
using the \Rfunction{[} sub-setting operator. Below, we mimic the behaviour of
\Rfunction{filterNA(, pNA = 0)} by calculating the row indices that should be removed, 
i.e. those that have at least on \Robject{NA} value and explicitly remove these row. 
This method allows one to devise and easily apply any filtering strategy.

<<removeNa, echo=TRUE, eval=FALSE>>=
whichRow <- which(is.na((qnt))) %% nrow(qnt)
qnt <- qnt[-whichRow, ]
@

See also the \Rfunction{plotNA} method to obtain a graphical overview of 
the completeness of a data set.

@@

<<make2exps2, echo=TRUE, cache=FALSE>>=
set.seed(1)
i <- sample(length(itraqdata), 35)
j <- sample(length(itraqdata), 35)
exp1 <- quantify(itraqdata[i], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
exp2 <- quantify(itraqdata[j], reporters = iTRAQ4,
                 parallel = FALSE, verbose = FALSE)
exp1 <- droplevels(exp1) 
exp2 <- droplevels(exp2)
table(featureNames(exp1) %in% featureNames(exp2))
exp1 <- combineFeatures(exp1,
                        groupBy = fData(exp1)$ProteinAccession)
exp2 <- combineFeatures(exp2,
                        groupBy = fData(exp2)$ProteinAccession)
head(featureNames(exp1))
head(featureNames(exp2))
@

The \Rfunction{droplevels} drops the unused \Robject{featureData} levels. 
This is required to avoid passing absent levels as \Robject{groupBy} 
in \Rfunction{combineFeatures}. 
Alternatively, one could also use \Robject{factor(fData(exp1)\$ProteinAccession)} 
as \Robject{groupBy} argument.

@@

\section{MS$^E$ data processing}\label{sec:mse}

\Rpackage{MSnbase} can also be used for MS$^E$ data independent acquisition from Waters instrument. 
The MS$^E$ pipeline depends on the Bioconductor \Rpackage{synapter} package \citep{synapter} that 
produces \Robject{MSnSet} instances for indvidual acquisitions. The \Rpackage{MSnbase} infrastructure can 
subsequently be used to further combine experiments, as shown in section \ref{sec:comb2} and 
apply \textit{top3} quantitation using the \Rfunction{topN} method.
